---
title: "Assignmet 4"
output: html_document
---

--
title: "Assignment 3"
author: "Linda Dumalo"
date: "October 1, 2019"
output:
  html_document:
    keep_md: yes
  pdf_document: default
--- 

```{r setup, include=FALSE, message=F}
library(gapminder)
library(tidyverse)
library(ggplot2)
library(knitr)
library(ggridges)
library(scales)
library(tidyr)
library(dplyr)
library(readr)
knitr::kable 
``` 

## Exercise 1: Univariate Data Reshaping 

#### Option 1

Make a tibble with one row per year, and columns for life expectancy for two or more countries.

```{r Option 1 Wide, include=TRUE}
# target table to have years on the left and "widen" country column 
OptI <- gapminder %>% 
  filter(country ==  "Canada" | country == "Mexico") %>% 
  select(country, year, lifeExp) %>% 
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = lifeExp)
OptI 
```

Relengthen the data. 

```{r}
OptI %>%
  pivot_longer(cols = c("Canada", "Mexico"), 
               names_to = "Country", 
               values_to = "Life_Exp") %>%
# Order wrong so rearrange 
arrange(Country, year, Life_Exp)
```

The Exercise asked to plot the widened data by plotting the 2 countries against each other. Interpreted this as plot one Country on x and 1 country on y. Where each point represents a year and the values for each country are a life epectancy. 

```{r}
OptI %>% 
  ggplot(aes(Canada, Mexico)) + 
  geom_point()
```
To compare though how the life expectancies differ between the countries per year it is easier to use the unwiden data 

```{r}
gapminder %>% 
  filter(country ==  "Canada" | country == "Mexico") %>% 
  select(country, year, lifeExp) %>% 
  ggplot(aes(year, lifeExp, fill = country, colour = country)) + 
  geom_point()
```


## **Multivariate Data** 
#### Option 1 

Make a tibble with one row per year, and columns for life expectancy and GDP per capita (or two other numeric variables) for Canada and Mexico. 


```{r Multivariatedata option1 v2}
# The observation is the combination of country, lifeExp and gdpPercap 
OptIM <- gapminder %>% 
  filter(country ==  "Canada" | country == "Mexico") %>% 
  select(year, country, lifeExp, gdpPercap) %>%
  pivot_wider(id_cols = year,
            names_from = country,
            names_sep = "_",
            values_from = c(lifeExp, gdpPercap))
OptIM
# is there a way to invert the order? ie have the column headers appear as Country_(lifeExp or gdppercap) 
```

Relengthen the data
Fuck. 

```{r}
OptIM %>%
  pivot_longer(cols = c(-year), # is this the correct line to start? I think so... 
               names_to = c(".value", "country"), 
               names_sep = "_") %>%
  arrange(country, year) 
# Data is too long and the country names are in the headers 
               
               
```

## **Exercise 3** - Table Joins 

```{r}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```
#### Exercise 3.1

Add an email column from the email tibble to the guest tibble.

```{r, Original email and guest tibbles}
head(guest)
head(email)
```

In order to perform the join functions, the email tibble format needs to be changed to match the guest tibble.

The changes that need to be made to the email tibble are: 

1. Have one observation ie name per row 
2. Change column name guest to name. 

```{r Step1}
email2 <- email %>% 
  separate_rows(guest, sep = ", ") %>% 
  rename("name" = "guest")
email2
```
```{r Step2} 
guest %>% 
  left_join(email2, by = "name") 
# left join was used to keep all the observations 
```

#### 3.2 

Who are there emails for but are not on the guest list? Not being on the guest list is to be intrepted as not having a meal or being in attendance to any event. 

```{r keep for now, discard upon revision}
# would think that you would want to do a right join in order to keep all observations in y ie the emails
guest %>% 
  right_join(email2, by = "name")
# This though is informing on who has emails and are on the guest list 
```
```{r, filtering for wedding attendance} 
guest %>% 
  full_join(email2, by = "name") %>% 
  filter(is.na(party)) %>% 
  select(party, name, attendance_wedding, email) 
```

Those who are not on the guest list but there was an email for were Turner Jones and Albert and Vivian Marshall. 

