---
title: "Assignment 2"
output: 
  html_document: 
    keep_md: true 
editor_options: 
  chunk_output_type: inline
---

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

```{r load packages, warning = FALSE, message = FALSE}
# load your packages here:
library(gapminder)
library(tidyverse)
``` 

### Exercise 1: Basic dplyr 

```{r} 
filter(gapminder, country %in%  c("Australia", "Canada", "Germany")) 
```

#### Exercise 1.2 

Use the pipe operator %>% to select “country” and “gdpPercap” from your filtered dataset in 1.1 

```{r} 
gapminder %>%
filter(country %in%  c("Australia", "Canada", "Germany")) %>% 
  select(country, gdpPercap) 
```

#### Exerciseilter( ) all countries that have experienced a decrease 
..... could it also be that when the instructions say  1.3 

Filter gapminder to all entries that have experienced a drop in life expectancy. Be sure to include a new variable that’s the increase in life expectancy in your tibble. Hint: you might find the lag() or diff() functions useful. 
``` {r} 

gapminder %>% 
  mutate(lifeExp, lag_lifeExp = lag(lifeExp)) %>% 
  mutate(lifeExp, delta_lifeExp = lifeExp - lag_lifeExp) %>%
  
```

Code chunk below will is to filter based on decreasing lifeExp ie negative values. 

``` {r}
gapminder %>% 
  mutate(lifeExp, lag_lifeExp = lag(lifeExp)) %>% 
  mutate(lifeExp, delta_lifeExp = lifeExp - lag_lifeExp) %>%
  filter(delta_lifeExp < 0) 
```
  Think however that this what is being asked is to : 
  have a column that says increase in life exptancy, sort based on decreasing life Exp 
  I also would like filter down to just the year 2007 (bonus though)
  