---
title: 'STAT 545A Homework Repository '
output:
  html_document: default
  pdf_document: default
theme: flatly
---

##### This repository (repo) is where assignments for **STAT 545A** are to be submitted. 

**STAT 545A is a course offered by the University of British Columbia to graduate students from a variety of disciplines to learn how to manipulate data with R.** 

For more inormation about the course go to the [Course website](https://stat545.stat.ubc.ca/)

My name is Linda Dumalo and I am currently doing a M.Sc of Biochemistry and Molecular Biology in Dr. Lindsay Eltis's lab. 

My research interests are: 

* Enzymology
 
  -Specifically, enzyme kinetics 

* Structural biology

* Microbiology

* Conducting proper experiments   ***controls, controls, controls***


Sometimes, science does not work or your data is less than stellar and not expected so a favourit quote from a fellow lab member is 

> *It is what it is* 


## **Going Macro:** Gapminder Exercise 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Gapminder is an important data frame since it looks at the following 6 variables for a country. The number of rows was 1706 which is definetly a lot more than the number of countries in the world! The reason for this is that the when gapminder is downloaded it is organized by country alphabetically and then by the year. 

```{r gapminder}
library(gapminder)
names(gapminder)
nrow(gapminder)
head(gapminder)
```

