---
title: 'STAT 545A Homework Repository '
output:
  html_document: default
  pdf_document: default
theme: flatly
---

##### This repository (repo) is where assignments for **STAT 545A** are to be submitted. 

**STAT 545A is a course offered by the University of British Columbia to graduate students from a variety of disciplines to learn how to manipulate data with R.** 

For more inormation about the course go to the [Course website](https://stat545.stat.ubc.ca/)

My name is Linda Dumalo and I am currently doing a M.Sc of Biochemistry and Molecular Biology in Dr. Lindsay Eltis's lab. 

My research interests are: 

* Enzymology
 
  -Specifically, enzyme kinetics 

* Structural biology

* Microbiology

* Conducting proper experiments   ***controls, controls, controls***


Sometimes, science does not work or your data is less than stellar and not expected so a favourit quote from a fellow lab member is 

> *It is what it is* 


## **Going Macro:** Gapminder Exercise 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Gapminder is an important data frame since it looks at the following 6 variables for a country:

```{r gapminder}
library(gapminder)
names(gapminder)
```
