---
title: "Assignment 2"
author: "Linda Dumalo"
date: "24/09/2019"
output:
  pdf_document: default
  html_document:
    keep_md: yes
--- 

```{r setup, include=FALSE, message=F}

library(gapminder)
library(tidyverse)
library(ggplot2)
library(knitr)
library(ggridges)
library(scales)
knitr::kable 
``` 


## Option 3 

Look at the spread of GDP per capita within the continents. 
Goal: Create a jitterplot with the continents on the x asis and the GDP per capita on the x 

```{r}
#maybe change some dimensions 
gapminder %>% 
  filter(year > 1980) %>% 
  # maybe rename the year variable to year here 
  rename("Year" = "year") %>% 
  ggplot(aes(continent, gdpPercap, colour = Year)) + 
  geom_jitter(width = 0.3) + 
  geom_point(alpha = 0.08) + 
  xlab("Continent") +
  scale_y_log10("GDP per Capita", labels= scales::dollar_format()) +
  theme_bw() 
  # bonus is to remove the grid lies 
```

## Task 5 Route 1 

Want to have the year on the x axis and on thee y ais lifeExp and each continent as a different trendline
Working backwards
need to take the average life expetancy for each continent per year and plot 
ideally would have a tibble that has 

Continent | year |  average lifeExp 

what if within ggplot you group by continent? 
```{r still a hot mess of lines} 
gapminder %>%
  ggplot(aes(year, lifeExp, group = continent)) +
  geom_path()
```

```{r, task 5, attempt 1}
gapminder %>%
  # group_by(year) %>%, <- include this and the output is the same 
  group_by(continent) %>%
  arrange(year) %>% 
  summarize(MeanLifeExp = mean(lifeExp)) 
  # the problem with summarizing seems to be that it obliderates the other variables since the error is saying that continent is not found 
  # option 1: find a way to summarize continent and meanlifeExp for each year and then pipe into ggplot 
  # ggplot(aes(x = (year it works but not with continent), MeanLifeExp))
  # this output would be the average lifeExp for all years and countries, not very useful 
``` 
```{r, task 5, attempt 3, from attempt 1}
# tried adding in the year, similar to what was done during class 
gapminder %>%
  # group_by(year) %>% 
  group_by(continent) %>%
  arrange(year) %>% 
  summarize(MeanLifeExp = mean(lifeExp), year = year[lifeExp = MeanLifeExp]) 
```

```{r, task 5, attempt 2}
#v1 
gapminder %>%
  group_by(year) %>% 
  arrange(year) %>% 
  summarize(MeanLifeExp = mean(lifeExp), continent = continent[lifeExp == MeanLifeExp]) 
  # error obtained: Column `continent` must be length 1 (a summary value), not 0 
```
```{r,  task 5, attempt 2}
# v2
gapminder %>%
    group_by(year, country) %>%  # having just country results in the same error as above 
  arrange(year) %>% 
  summarize(MeanLifeExp = mean(lifeExp), continent = continent[lifeExp == MeanLifeExp]) 
```

```{r,  task 5, attempt 2}
# v3 Not sure what I am trying... 
gapminder %>%
  group_by(year, country) %>%  # having just country results in the same error as above 
  arrange(year) %>% 
  summarize(MeanLifeExp = mean(lifeExp), continent = continent[lifeExp == MeanLifeExp]) %>% 
  ungroup() %>% 
  group_by(continent) 
```



```{r task 5, attempt 4} 
# v1 
# grouped by country but then lost the data for the year 
gapminder %>%
  group_by(country) %>%
  summarize(MeanLifeExp = mean(lifeExp), year = year[lifeExp = MeanLifeExp], continent = continent[lifeExp = MeanLifeExp]) 
```

```{r task 5, attempt 4} 
# v2 
# FAIL  
gapminder %>%
  summarize(MeanLifeExp = mean(lifeExp), year = year[lifeExp = MeanLifeExp], continent = continent[lifeExp = MeanLifeExp]) 
```

## Task 5 Route 2 

Create a histogram of the difference in life expectancy between 1977 to 2007 

Use lag and stuff 

```{r}
gapminder %>% 
  filter(year %in% c(1977, 2007)) %>% # this is not filtering to 1977 or 2007 tho, 
  # not grouping by continent first 
  group_by(country) %>% 
  arrange(year) %>% 
  # ^ precaution 
  mutate(DeltaLifeExp = difference(lifeExp)) %>% 
  drop_na() %>% 
  ungroup() %>%
  group_by(continent) %>%
  arrange(continent) %>% 
  summarise(MeanDeltaLifeExp = mean(DeltaLifeExp)) %>% 
  ggplot(aes(continent, MeanDeltaLifeExp)) + 
  geom_bar() 
```

```{r}
gapminder %>% 
  filter(year %in% c(1977, 2007)) %>% # this is not filtering to 1977 or 2007 tho, 
  # not grouping by continent first 
  group_by(country) %>% 
  arrange(year) %>% 
  # ^ precaution 
  mutate(DeltaLifeExp = difference(lifeExp)) %>% 
  drop_na() %>% 
  ungroup() %>%
  group_by(continent) %>%
  arrange(continent) %>% 
  summarise(MeanDeltaLifeExp = mean(DeltaLifeExp) %>%
  ggplot(aes(continent, DeltaLifeExp)) + 
  geom_bar() 
```

