---
title: "Untitled"
output: html_document
---

## Task 5 Route 1 

Want to have the year on the x axis and on thee y ais lifeExp and each continent as a different trendline
Working backwards
need to take the average life expetancy for each continent per year and plot 
ideally would have a tibble that has 

Plot the life expectancy in geom_path per year and have  different lines each representing the continent 
My initial approach (approach 1) involved summarizing the gapminder data to have the average life expectancy for each year for each continent using group_by() and summarize() 
So the table headings might read something like 
Continent | Year | Average LifeExp 
Asia  | 1952 | x |
Asia | 1957 | y | 
The reason that I wanted to do this is because I recall in class that ggplot will not recognize any grouping in previous pipes. 
However, I am having a hard time generating this table  

Continent | year |  average lifeExp 

what if within ggplot you group by continent? 

```{r still a hot mess of lines} 
gapminder %>%
  ggplot(aes(year, lifeExp)) +
  geom_path()
```

```{r, task 5, attempt 1}
gapminder %>%
  # group_by(year) %>%, <- include this and the output is the same 
  group_by(continent) %>%
  arrange(year) %>% 
  summarize(MeanLifeExp = mean(lifeExp)) 
  # the problem with summarizing seems to be that it obliderates the other variables since the error is saying that continent is not found 
  # option 1: find a way to summarize continent and meanlifeExp for each year and then pipe into ggplot 
  # ggplot(aes(x = (year it works but not with continent), MeanLifeExp))
  # this output would be the average lifeExp for all years and countries, not very useful 
``` 
```{r, task 5, attempt 3, from attempt 1}
# tried adding in the year, similar to what was done during class 
gapminder %>%
  # group_by(year) %>% 
  group_by(continent) %>%
  arrange(year) %>% 
  summarize(MeanLifeExp = mean(lifeExp), year = year[lifeExp = MeanLifeExp]) 
```

```{r, task 5, attempt 2}
#v1 
gapminder %>%
  group_by(year) %>% 
  arrange(year) %>% 
  summarize(MeanLifeExp = mean(lifeExp))
  # error obtained: Column `continent` must be length 1 (a summary value), not 0 
```
```{r,  task 5, attempt 2}
# v2
gapminder %>%
    group_by(year, country) %>%  # having just country results in the same error as above 
  arrange(year) %>% 
  summarize(MeanLifeExp = mean(lifeExp), continent = continent[lifeExp == MeanLifeExp]) 
```

```{r,  task 5, attempt 2}
# v3 Not sure what I am trying... 
gapminder %>%
  group_by(country, continent) %>%  # having just country results in the same error as above 
  arrange(year) %>% 
  summarize(MeanLifeExp = mean(lifeExp)) 
```

```{r,  task 5, attempt 2}
# v4 Not sure what I am trying... 
gapminder %>%
  group_by(country, continent) %>%  # having just country results in the same error as above 
  arrange(year) %>% 
  # summarize(MeanLifeExp = mean(lifeExp)) %>% 
  ggplot(aes(year, lifeExp)) + 
  geom_path()
```

```{r task 5, attempt 4} 
# v1 
# grouped by country but then lost the data for the year 
gapminder %>%
  group_by(country) %>%
  summarize(MeanLifeExp = mean(lifeExp), year = year[lifeExp = MeanLifeExp], continent = continent[lifeExp = MeanLifeExp]) 
```

```{r task 5, attempt 4} 
# v2 
# FAIL  
gapminder %>%
  summarize(MeanLifeExp = mean(lifeExp), year = year[lifeExp = MeanLifeExp], continent = continent[lifeExp = MeanLifeExp]) 